\chapter{Conclusions and future work}\label{chapter:conclusions}


\subsubsection{Auralization paper}\todo{rewrite and merge}
In order to investigate the impact of aircraft sound on humans a tool was
developed to auralise aircraft sound. The goal was to develop a tool that can
create auralisations that sound similar to the target aircraft and is capable of
creating auralisations of the current fleet of aircraft.

A propagation model based on geometrical acoustics was developed and the
emission synthesis was based on features extracted from recordings with the use
of an inverse propagation model and a feature extraction algorithm. A listening
test was conducted to determine whether the chosen features are sufficient for
creating auralisations that sound similar to the actual aircraft.

From the listening test it follows that listeners can discriminate between
sounds of different aircraft types, in case of both recordings and
auralisations. The similarity ratings between recordings and auralisations are
similar to the similarity ratings between the two aircraft type, implying the
auralisations sound like different aircraft than those that were recorded.

Therefore, future work should emphasise on further improving the simulations.
The authors think that the most important point to improve is the simulation of
the blade passing frequency and harmonics, and that would require a better
estimation of their powers and bandwidths. Aside from that, it would seem the
auralisation tool is capable of delivering plausible auralisations.

\subsubsection{Turbulence paper}\todo{rewrite and merge}
Fluctuations in the refractive-index field due to variations in temperature and
wind affects sound propagation and causes audible modulations. A method was
presented for generating sequences of modulations and applying these to
monochromatic as well as broadband signals.

A Rytov approximation to first-order refractive-index flucutuations results in a complex
phase which we can write as a log-amplitude $\chi$ and phase $S$ fluctuation.
The propagating sound is modelled as a time-varying channel where we consider
two sequences, one for the log-amplitude fluctuations, and another for the phase
fluctuations.

The fluctuations are frequency-dependent and therefore a filter was designed to take that into account.
% A Gaussian applied filter was used to model the atmospheric turbulence because of its simplicity and because it is computationally least demanding.
A Gaussian turbulence spectrum was considered, but the general method can be
used with other turbulence spectra as well. The Von Karman spectrum describes
real turbulence spectra typically better than the Gaussian spectrum, however,
the Von Karman spectrum is computationally much more demanding.
% The Gaussian model also allows implementing the phase fluctuation with a Variable Delay Line (VDL).

Examples are shown where the method is applied to a tone and to an aircraft
auralization. The aircraft auralization spectrogram shows several spikes
corresponding to amplitude modulations as well as an increase in the amount of
decorrelation. Furthermore the transverse velocity dependence on the frequency
content of the modulations is demonstrated.

According to the authors the method results in more realistic sounding
auralizations, but this has not been validated yet with listening tests. The
implementation of the model that was used in this paper to generate the figures
can be found at \cite{Rietdijk2016}.



\section{Future work}


\subsection{Develop better indicators for annoyance and sleep disturbance}
