The research leading to these results has received funding from
the People Programme (Marie Curie Actions) of the European Union's Seventh
Framework Programme FP7/2007-2013 under REA grant agreement number 290110,
SONORUS "Urban Sound Planner".

Four years ago I began to work on aircraft auralisation at Empa in Switzerland.
Not yet fully grasping the importance of the aircraft noise issue I soon found
out, enjoying aircraft sounds daily, both
at work, and at home! So, thank you to everyone flying in and out of Z\"{u}rich
Airport during that period for the wonderful learning experience!

Thankfully, there was far more than just the annoyance and sleep disturance due to
these aircraft. There were my colleagues at Empa, whom I thank for the
discussions and the introduction to your secret language and its
dialects, some of which were just impossible to understand. You
tried hard and I appreciate the attempts, but I think its safe to say I'll never
speak \say{Switzerdeutsch}. What I will remember is the \say{li} at the end of
words, since it does sound funny :)

Kurt H., I am glad I had you as a supervisor during these years at Empa. I've
learned a lot from you and from our discussions. It was very nice being able to
just walk in, discuss, and write down our ideas. We tried real hard getting to
the bottom of that pile of old seminar abstracts! Reto, thank you also for the
useful discussions, for contributing some decent music to the playlist, and for
ensuring the music was on by the time I got in. Christoph, thank you for helping
with the aircraft and airport \say{things}, and for providing that mountain of
data. It was good to be able to discuss and compare our approaches.

Then, there are my colleagues in the SONORUS project. Four years ago we embarked
on an \say{Urban Sound Planning} journey and faced soundwalks, noise maps,
discussions, presentations, and pubs. And, indeed, our research projects.
Frustrating at times, I am still glad for the experience and our time together.

Time to fly to Sweden: Chalmers. Wolfgang, thank you for the supervision. We did
not meet often, and mails occasionally ended up in a black hole, but the meetings we
had, and the feedback I got, was very useful, thank you! Jens, I recall you said
you \say{had an idea} on auralisation and turbulence. I don't think I could have
imagined at the time what kind of literature would follow. Thank you for that
initial idea, and thank you for the assistence along the way. Finally, I would
like to say thank you to everyone at the division for the talks, discussions,
fikor, puzzle tour(s) and the \say{gezellige} environment!


% % TODO is that also required here?
%
% \vskip 2pc
%
% \noindent \thesisauthor
%
% \noindent \thesiscity, August\  2017  % Since dedication is written a month or more before the actual thesis date, \thesismonth and \thesisyear is not used here.
